Semantic Preserving Bijective Mappings for Expressions involving Special Functions between Computer Algebra Systems and Document Preparation Systems

### 2018-11-30 main.tex
Fixed Fig4 (close #21)
Fix get
Closes #29
Reword comment on generic LaTeX
Close #32

\section{Related Work}\label{sec:related-work}
Since \LaTeX{} became the de facto standard for writing papers in mathematics, most \gls*{cas} provide simple functions to import and export mathematical \LaTeX{} expressions\footnote{The selected \gls*{cas} Maple, Mathematica, Matlab, and SageMath provide import and/or export functions for \LaTeX:\\
Maple, \url{http://www.maplesoft.com/support/help/Maple/view.aspx?path=latex} seen 06/2017;
Mathematica, \url{https://reference.wolfram.com/language/tutorial/GeneratingAndImportingTeX.html} seen 06/2017;
Matlab, \url{https://www.mathworks.com/help/symbolic/latex.html} seen 06/2017;
SageMath, \url{http://doc.sagemath.org/html/en/tutorial/latex.html} seen 06/2017.}. Those tools have two essential problems. They are only able to import simple mathematical expressions, where the semantics are unique. For example, the internal \LaTeX{} macro \verb|\frac| always indicates a fraction. However, for more complex expressions, e.g., the Jacobi polynomial in \Cref{tab:JacobiP-usecase}, the import functions fail. The second problem appears in the export tools. Mathematical expressions in \gls*{cas} are fully semantic. Otherwise the \gls*{cas} wouldn't be able to compute or evaluate the expressions. During the export process, the semantic information gets lost, because generic \LaTeX{} is not able to carry sufficient semantic information. Because of these problems, an exported expression cannot be imported to the same system again in most cases (except for simple expressions such as those described above). Our tool attempts to solve these problems and provide round-trip translations between \LaTeX{} and \gls*{cas}.

@@

\begin{wrapfigure}[10]{r}{0.38\textwidth} % optional parameter gives "# of narrow lines", very useful! 
\vspace{4pt}
\begin{minipage}{0.38\textwidth}
\center
\begin{tabular}{lc}
\hline
Notation & Expression \\
\hline
Infix & $(a+b) \cdot x$\\
Prefix & $\cdot + a\ b\ x$\\
Postfix & $a\ b + x\ \cdot$\\
Functional & $\cdot(+(a, b), x)$\\
\hline
\end{tabular}
%\vspace{-5pt}
\caption{The mathematical expression `$(a+b) \cdot x$' in infix, prefix, postfix and functional notation.}
\label{tab:notations}
\end{minipage}
%\vspace{-5pt}
\end{wrapfigure}

@@

@@

The translator was designed to be easily extendable. This allows one to implement translations for other \gls*{cas} without much effort.
An extensive weakness of the translator is that the input needs to be semantically enhanced, in this case by the special macros from the \gls*{dlmf}.
Therefore, without exclusively given semantic information, the translator is not able to translate functions. Currently, we are working on mathematical information retrieval techniques which will allow for an extension of the translator to generic \LaTeX{} inputs.

>>>>

\section{Related Work}\label{sec:related-work}
Since \LaTeX{} became the de facto standard for writing papers in mathematics, most \gls*{cas} provide simple functions to import and export mathematical \LaTeX{} expressions\footnote{The selected \gls*{cas} Maple, Mathematica, Matlab, and SageMath provide import and/or export functions for \LaTeX:\\
Maple, \url{http://www.maplesoft.com/support/help/Maple/view.aspx?path=latex} seen 06/2017;
Mathematica, \url{https://reference.wolfram.com/language/tutorial/GeneratingAndImportingTeX.html} seen 06/2017;
Matlab, \url{https://www.mathworks.com/help/symbolic/latex.html} seen 06/2017;
SageMath, \url{http://doc.sagemath.org/html/en/tutorial/latex.html} seen 06/2017.}. Those tools have two essential problems. They are only able to import simple mathematical expressions, where the semantics are unique. For example, the internal \LaTeX{} macro \verb|\frac| always indicates a fraction. However, for more complex expressions, e.g., the Jacobi polynomial in \Cref{tab:JacobiP-usecase}, the import functions fail. The second problem appears in the export tools. Mathematical expressions in \gls*{cas} are fully semantic. Otherwise the \gls*{cas} wouldn't be able to compute or evaluate the expressions. During the export process, the semantic information is lost, because generic \LaTeX{} is not able to carry sufficient semantic information. Because of these problems, an exported expression cannot be imported to the same system again in most cases (except for simple expressions such as those described above). Our tool attempts to solve these problems and provide round-trip translations between \LaTeX{} and \gls*{cas}.

@@

@@

\begin{wrapfigure}[10]{r}{0.38\textwidth} % optional parameter gives "# of narrow lines", very useful! 
\vspace{-5pt}
\begin{minipage}{0.38\textwidth}
\center
\begin{tabular}{lc}
\hline
Notation & Expression \\
\hline
Infix & $(a+b) \cdot x$\\
Prefix & $\cdot + a\ b\ x$\\
Postfix & $a\ b + x\ \cdot$\\
Functional & $\cdot(+(a, b), x)$\\
\hline
\end{tabular}
%\vspace{-5pt}
\caption{The mathematical expression `$(a+b) \cdot x$' in infix, prefix, postfix and functional notation.}
\label{tab:notations}
\end{minipage}
\vspace{3pt}
\end{wrapfigure}

@@

The translator was designed to be easily extendable. This allows one to implement translations for other \gls*{cas} without much effort.
However, most \LaTeX{} sources, such as arXiv, are given in generic \LaTeX{}. Semantic \LaTeX{}, which is a prerequsite for our translator, is currently prevalent in the DLMF and DRMF projects alone. Therefore, without exclusively given semantic information, the translator is not able to translate functions. Currently, we are working on mathematical information retrieval techniques which will allow for an extension of the translator to generic \LaTeX{} inputs.
